---
title: "individual2"
author: "Guang Yi Lim"
date: "February 26, 2019"
output: html_document
---

```{r setup, include=FALSE}
data(happy,package ="productplots")
HAPPY <-readRDS("HAPPY.rds")
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```
Part 1
1.Data cleaning: the values "IAP", "DK" and "NA" all encode missing values. Replace all of these instances by the value NA.
```{r}
  HAPPY[HAPPY =='IAP' | HAPPY =='DK' |HAPPY =='NA']<-NA
```
</b>
2. Check the type of the variable and cast into the right type (factor variable for categorical variables). For age, change "89 OR OLDER" to 89 and assume the variable should be numeric. (Hint: see ?as.character and ?as.numeric)
```{r}
  HAPPY$HAPPY <-as.factor(HAPPY$HAPPY)
  HAPPY$AGE[HAPPY$AGE =="89 OR OLDER"]<-89
  HAPPY$AGE <- as.numeric(HAPPY$AGE)
  HAPPY$SEX <-as.factor(HAPPY$SEX)
  HAPPY$MARITAL <-as.factor(HAPPY$MARITAL)
  HAPPY$DEGREE <-as.factor(HAPPY$DEGREE)
  HAPPY$FINRELA <-as.factor(HAPPY$FINRELA)
  HAPPY$HEALTH <-as.factor(HAPPY$HEALTH)
  HAPPY$WTSSALL <-as.numeric(HAPPY$WTSSALL)
  HAPPY$PARTYID <-as.factor(HAPPY$PARTYID)
```
3. Factor the data
```{r}
HAPPY$HEALTH<- factor(HAPPY$HEALTH,c("POOR","FAIR","GOOD","EXCELLENT","DK"))
HAPPY$FINRELA <-factor(HAPPY$FINRELA,c("FAR BELOW AVERAGE","BELOW AVERAGE","AVERAGE","ABOVE AVERAGE","FAR ABOVE AVERAGE" ))
HAPPY$MARITAL <- factor(HAPPY$MARITAL ,c("NEVER MARRIED","MARRIED","DIVORCED","SEPERATED","WIDOWED"))
HAPPY$DEGREE <-factor(HAPPY$DEGREE,c("HIGH SCHOOL","LT HIGH SCHOOL","JUNIOR COLLEGE" ,"BACHELOR","GRADUATE","DK"))
HAPPY$PARTYID<-factor(HAPPY$PARTYID,c("IND,NEAR REP","INDEPENDENT","IND,NEAR DEM","NOT STR DEMOCRAT","STRONG DEMOCRAT","STRONG REPUBLICAN","NOT STR REPUBLICAN", "OTHER PARTY"))
```

Part 2
1.
```{r}
temp <-HAPPY[!is.na(HAPPY$HAPPY),]
ggplot(HAPPY,aes(x=YEAR,fill =SEX))+geom_histogram(bindwidth=0.5,na.rm=TRUE)+facet_wrap(~SEX)

```
</b>from the graph we can see the data is distributed more towards women than men
2.
```{r} 

ggplot(temp,aes(x=HAPPY,fill =SEX))+geom_bar(na.rm =TRUE)+coord_flip() +facet_wrap(~MARITAL+SEX)
```

